Adaptive broadcast tree for multicores

### 2013-05-06 paper.tex
Simulation for ring-based outer communication

@@

First, we discuss what kind of topologies we use to communicate
between nodes. We evaluate tree based implementations. Other
potentially interesting topologies are rings.

@@

@@

@@

\begin{figure}[htb]
  \centering
  \begin{tikzpicture}[scale=.35,transform shape]
    % Insert visualization
    \input{visu_mst_sort}
  \end{tikzpicture}
  \caption{Visualization of operations for sending a message along the
    tree. Red boxes represent send operations, blue boxes represent
    receive operations and arrows messages sent between cores. After
    running the MST algorithm, we optimize scheduling in every node to
    send messages to in decreasing order of their latency.}
  \label{fig:mst_gruyere_operations_sorted}
\end{figure}

@@

\begin{table}[htb]
  \centering
  \begin{tabular}{llll}
    \toprule
    topology & scheduling & time & factor \\
    \midrule
    MST & naive                     & 241 & 1.00 \\
    MST & highest link weight first & 151 & 0.63 \\
    MST & longest path first        & 151 & 0.63 \\
    \bottomrule
  \end{tabular}
  \caption{Simulation results for simplified machine model}
  \label{tab:sim_results}
\end{table}

@@

>>>>

% keywords: modeling, simulation, overlay networks, scheduling

@@

First, we discuss what kind of topologies we use to communicate
between nodes. We evaluate tree based implementations: binary trees,
minimum spanning trees, hierarchical tress and clustering
techniques. Other potentially interesting topologies are rings.

@@

%-------------------------------------------------
\subsubsection{Hierarchical}

We now consider NUMA node for coming up with a broadcast tree. We
select one coordinator node in every node and build a binary tree
for the coordinators. We call this topology connecting nodes across
NUMA domains the \emph{outer} network.

Each coordinator is also responsible for sending a message to all
other cores of the node. We currently do this sequentially, since the
number of cores in a node is typically very small. Alternatively, a
binary tree could be used as well. We call overlay networks within
NUMA nodes \emph{inner} networks.

We show an example hierarchical tree in
Figure~\ref{fig:gruyere_hierarchy}.

\begin{figure}
\begin{tikzpicture}[>=latex,line join=bevel,scale=.5]
  \pgfsetlinewidth{.5bp}
\input{figs/final_gruyere_hierarchical}
\end{tikzpicture}
\caption{Hierarchical tree for a 8x4x1 multicore}
\label{fig:gruyere_hierarchy}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Ring topology}

Next, we evaluate ring topologies. Rings have the advantage that the
implicitly acknowledge messages once they arrive back to the
sender. This is beneficial for some algorithms such as quorums, where
nodes report back to the sender.

The downside is that some state is required on every node to remember
whether or not a message was received before. This however only works
if forwarding along the ring is only done after all local nodes
acknowledge a message. If messages are just forwarded as they arrive,
message transport does never terminate.

To add more parallelism, we add a second ring, which we send the
message on in parallel. The overlay ring network is visualized in
Figure~\ref{fig:gruyere_ring}.

\begin{figure}
\begin{tikzpicture}[>=latex,line join=bevel,scale=.5]
  \pgfsetlinewidth{.5bp}
\input{graphs/final_ring}
\end{tikzpicture}
\caption{Outer overlay as ring with sequential inner overlay for a
  8x4x1 multicore}
\label{fig:gruyere_ring}
\end{figure}

@@

\begin{figure}[htb]
  \centering
  \begin{tikzpicture}[scale=.35,transform shape]
    % Insert visualization
    \input{visu_mst_sort}
  \end{tikzpicture}
  \caption{Visualization of operations for sending a message along the
    tree. Red boxes represent send operations, blue boxes represent
    receive operations and arrows messages sent between cores. After
    running the MST algorithm, we optimize scheduling in every node to
    send messages in decreasing order of their latency.}
  \label{fig:mst_gruyere_operations_sorted}
\end{figure}

@@

\begin{table}[htb]
  \centering
  \begin{tabular}{lllll}
    \toprule
    topology & scheduling & time & factor & figure \\
    \midrule
    MST & naive                     & 241 & 1.00 & 
        \ref{fig:mst_gruyere_operations} \\
    MST & highest link weight first & 151 & 0.63 & 
        \ref{fig:mst_gruyere_operations_sorted} \\
    MST & longest path first        & 151 & 0.63 & 
        \stefan{? AFAIK, same as
        \ref{fig:mst_gruyere_operations_sorted} } \\
    hierarchical & highest link weight first & 151 & 0.63 &
        \ref{fig:visu_hierarchy} \\
    ring & highest link weight first & 181 & 0.75 &
        \ref{fig:visu_ring} \\
    \bottomrule
  \end{tabular}
  \caption{Simulation results for simplified machine model}
  \label{tab:sim_results}
\end{table}

@@

\begin{figure}[htb]
  \centering
  \begin{tikzpicture}[scale=.35,transform shape]
    % Insert visualization
    \input{visu_hierarchy}
  \end{tikzpicture}
  \caption{Visualization of hierarchical tree broadcast. A binary tree
    spans across NUMA nodes followed by sequential send operations
    within the NUMA node. %
    Red boxes represent send operations, blue boxes represent receive
    operations and arrows messages sent between cores.  }
  \label{fig:visu_hierarchy}
\end{figure}

\begin{figure}[htb]
  \centering
  \begin{tikzpicture}[scale=.35,transform shape]
    % Insert visualization
    \input{visu_ring}
  \end{tikzpicture}
  \caption{Visualization of ring based broadcast. Two rings
    spans across NUMA nodes followed by sequential send operations
    within the NUMA node. %
    Red boxes represent send operations, blue boxes represent receive
    operations and arrows messages sent between
    cores. \stefan{Scheduling is broken. If we want acks, we can only
      forward a message along the ring when all the local nodes
      acknowledge a message}}
  \label{fig:visu_ring}
\end{figure}
